paper1_figures

### 2023-09-14 paper1_markdown.Rmd
14.09

@Torsvik2011: in the introduction says that people in social dilemma are confused and can make random choices = argument against the completely rational human (see wich papers they cited)

@@

@@

a.  Climate change as social dilemmas
b.  Definition of commons  commons as social dilemmas
c.  Article that says that social dilemmas are not the same as commons (Cole & Grossman, 2010)
d.  But: Reindl et al. (2019) and Rommel et al. (2015)  behaviour in social dilemmas is informative about behaviours in commons + @Balliet2011: in the introduction (pages 1-2) there is a good explanation of social dilemma's utility to investigate cooperation
e.  Existing theory about the commons = similar structure than for social dilemmas: rational human beings are supposed to fail (TOC, Hardin, 1986), but in reality cooperation is often observed (for example Ostrom, 1990 + find others)

@@

Social identity have a consensual positive effect on cooperation: the more individuals feel a strong identification with a group, the more they cooperate with other members of the group (e.g. @Arora2016; @Noonan2016; @Gomez-Ruiz2020). For example, in a field study about community-organized dog parks in the United States, @Noonan2016 found that a strong sense of community was related to a higher amount of voluntary activities to maintain the parks in good conditions. This is explained by the fact that, when people strongly identify with a social group, they are more concerned about their group's well-being and reputation [@VanVugt2009] and they accord more importance to group-related goal [@Arora2016]. A strong social identity also maintains a high level of cooperation over time [@Zhang2019]. Moreover, independently of the strength of group identification, people cooperate more with other people in the same group than with out-group people [@Dorrough2015].

However, the fact that not all members identify with their group to the same extent can lead to inequalities on cooperation. For example, @VanVugt2009 mentioned that high identified individual can even compensate, by cooperating more, for other in-group members' defection. The same result was also found by @Arora2016: high identified individuals cooperated more than low identified ones when faced to an in-group participant who defected (this difference was not found when the defector was not a group member). However, this compensation came with a cost, since participants who compensated for an in-group defector reported lower trust toward this person.

Another well replicated result is the positive effect of intergroup competition on intragroup cooperation. In fact, in-group identification can be enhanced via a competition between different groups because in such a context individual's and group's interests are aligned [@Puurtinen2009]. In a social dilemma experiment, these authors demonstrated that cooperation was higher in an intergroup-competition condition than in a non-competitive condition. Moreover, when participants had to compete against other groups, they reported higher levels of anger toward in-group free-riders and they felt more guilty after a defection decision. In another experiment, intergroup competition reduced resources consumption (and therefore enhanced cooperation), while individual, dispositional competitiveness predicted an higher level of harvesting [@Nockur2020]. This shows that the increase of cooperation via competition is due to group membership factors, and not to competition per se.

However, despite the observed beneficial effects of inducing an intergroup competition, in a real-life situation this could be harmful. This is especially true in case of resources that are shared between several groups, which are at greater risk of over-exploitation in case of intergroup competition [@VanVugt2009]. A possible solution to maintain the positive effects of group identification, while eliminating the risk of over-exploitation caused by intergroup competition, is to prime a superordinate identity [@VanVugt2009]. In fact, intergroup cooperation can be enhanced by making salient the common characteristics of the different groups, instead of their differences [@Kramer1984]. Another effective way to create a superordinate identity is to induce the feeling of a common fate, for example through a collective risk. In fact, a common fate allows a group to become real [@Zhang2019], and can therefore erase intergroup competition.

@@

In general, social identity seems to have a positive and largely observed effect on individuals' cooperation. This is due to the fact that humans have a very developed sense of belonging to social groups, and that this has an impact on their willingness to help other members of their group [@VanVugt]. Since priming a common social identity is quite simple (e.g. @Kramer1984; @Zhang2019), this could be a very powerful method to enhance cooperation in an easy and cheap way.

@@

>>>>

@Torsvik2011: in the introduction says that people in social dilemma are confused and can make random choices = argument against the completely rational human (see wich papers they cited) <doi:10.1098/rstb.2010.0135>: in the introduction they cite papers that have observed cooperation

@@

-   @VanLange1998: many behaviours that tend to serve our personal well-being are detrimental to the environment

@@

a.  Climate change as social dilemmas
b.  Definition of commons  commons as social dilemmas
c.  Article that says that social dilemmas are not the same as commons (Cole & Grossman, 2010)
d.  But: Reindl et al. (2019) and Rommel et al. (2015)  behaviour in social dilemmas is informative about behaviours in commons + @Balliet2011: in the introduction (pages 1-2) there is a good explanation of social dilemma's utility to investigate cooperation

-   @Sussman2016: at page 11 mention that simulation of commons can be realistic why people feel real emotions (see the articles they cited)

e.  Existing theory about the commons = similar structure than for social dilemmas: rational human beings are supposed to fail (TOC, Hardin, 1986), but in reality cooperation is often observed (for example Ostrom, 1990 + find others)

@@

Social identity have a consensual positive effect on cooperation: the more individuals feel a strong identification with a group, the more they cooperate with other members of the group (e.g. @Arora2016; @Noonan2016; @Gomez-Ruiz2020). For example, in a field study about community-organized dog parks in the United States, @Noonan2016 found that a strong sense of community was related to a higher amount of voluntary activities to maintain the parks in good conditions. This is explained by the fact that, when people strongly identify with a social group, they are more concerned about their group's well-being and reputation [@VanVugt2009] and they accord more importance to group-related goal [@Arora2016]. A strong social identity also maintains a high level of cooperation over time [@Zhang2019]. Moreover, the link between group identification and cooperation is the same for individualist and collectivist people [@Chen2007], which is a further confirmation of the strength of this effect. Finally, independently of the strength of group identification, people cooperate more with other people in the same group than with out-group people [@Dorrough2015; @Aksoy2019].

However, the fact that not all members identify with their group to the same extent can lead to inequalities on cooperation. For example, @VanVugt2009 mentioned that high identified individual can even compensate, by cooperating more, for other in-group members' defection. The same result was also found by @Arora2016: high identified individuals cooperated more than low identified ones when faced to an in-group participant who defected (and this difference was not found when the defector was not a group member). However, this compensation came with a cost, since participants who compensated for an in-group defector reported lower trust toward this person.

Another well replicated result is the positive effect of intergroup competition on intragroup cooperation. In fact, in-group identification can be enhanced via a competition between different groups, because in such a context individual's and group's interests are aligned [@Puurtinen2009]. In a social dilemma experiment, these authors demonstrated that cooperation was higher in an intergroup-competition condition than in a non-competitive condition. Moreover, when participants had to compete against other groups, they reported higher levels of anger toward in-group free-riders and they felt more guilty after a defection decision. In another experiment using a public resource game, intergroup competition reduced resources consumption (and therefore enhanced cooperation), while individual, dispositional competitiveness predicted an higher level of harvesting [@Nockur2020]. This shows that the increase of cooperation via competition is due to group membership factors, and not to competition per se.

However, despite the observed beneficial effects of inducing an intergroup competition, in a real-life situation this could be harmful. This is especially true in case of resources that are shared between several groups, which are at greater risk of over-exploitation in case of intergroup competition [@VanVugt2009]. A possible solution to maintain the positive effects of group identification, while eliminating the risk of over-exploitation caused by the competitive context, is to prime a superordinate identity [@VanVugt2009]. In fact, intergroup cooperation can be enhanced by making salient the common characteristics of the different groups, instead of their differences [@Kramer1984]. Another effective way to create a superordinate identity is to induce the feeling of a common fate, for example through a collective risk. In fact, a common fate allows a group to become real [@Zhang2019], and can therefore erase intergroup competition.

@@

In general, social identity seems to have a positive and largely observed effect on individuals' cooperation. This is due to the fact that humans have a very developed sense of belonging to social groups, and that this has an impact on their willingness to help other members of their group [@VanVugt2009]. Since priming a common social identity is quite simple (e.g. @Kramer1984; @Zhang2019), this could be a very powerful technique to enhance cooperation in an easy and cheap way.

## Values and personality traits

Personal values and personality traits can play an important role in predicting cooperative behaviour [@Hilbig2018]. One of the most studied is the Social Value Orientation (SVO), that divides people in pro-social and pro-self based on how much they value their own and others' outcomes in situations of interdependence [@Balliet2009]. The effect of SVO on cooperation is very consensual: in a meta-analysis, 81 out of the 82 papers examined reported than pro-social cooperated significantly more that pro-self [@Balliet2009].

However, this general effect is subjected to several moderators. For example, pro-self individuals cooperated more in a game where incentives to cooperation were present than in a game without such incentives, while pro-social did not displayed this difference and showed high levels of cooperation in both contexts [@Emonds2011]. A similar effect is also described by @Balliet2009, who found a stronger effect of SVO in unpaid games than in paid ones, because, when monetary rewards were present, pro-self increased their cooperation. Therefore, pro-self individuals also cooperate, but only when cooperation serves their personal interest. In fact, in a study about commuting habits, the extent to which the cooperative behaviour (i.e. choosing to commute by public transportation or carpooling) was perceived as efficient impacted pro-self individual's decision more than pro-social's [@VanLange1998].

Moreover, SVO has a stronger effect if there is an uncertainty about the number of individuals having access to a common resource: when the number of other participants was not known, pro-social, but not pro-self, reduced their harvesting comparing to a group-size certainty condition [@DeKwaadsteniet2008]. The same authors demonstrated that this result is explained by the fact that pro-social individuals want to maximize the joint outcome, while pro-self only think at their own personal interest. Finally, SVO effect is partly explained by trust, i.e. the extent to which people think that their interaction partners will also cooperate. In fact, in @VanLange1998, pro-social individuals with a low level of trust did not make different decisions than pro-self individuals, while the most cooperative were pro-social with an high level of trust.

Other studies also focused on how much people are self- or social-oriented by using other constructs. For example, @Liu2009a demonstrated that priming a common fate to induce a context of interdependence significantly increased the cooperation level of participants with an independent self, but had no effect on participants with an interdependent self (who already displayed high levels of cooperation). Moreover, in a social dilemma experiment, competitive, individualistic people were motivated just by personal gain, while people who saw themselves as a part of a group were more interested on maximizing the income of each member, even at the cost to have a smaller personal payoff [@Probst1999]. Also, high allocentric (i.e collectivist) individuals cooperated more than low allocentric ones when the perceived social norms are cooperative: this means that they are more sensitive to social cues and therefore they adapt their behaviour to social norms [@Chen2007].

Finally, there are other personal values and traits that can affect cooperation. For example, pro-environmental values are negatively correlated with harvesting from a common resource, but only when the resource is scarce; therefore these values are more likely to motivate difficult, rather than easy, ecological behaviours [@Sussman2016]. Levels of harvesting from a common resource are also lower between left-wingers than right-wingers [@Fosgaard2019]. Moreover, humility and honesty are significant predictor of cooperation [@Kieslich2014], and this is due to the fact that people with such characteristics manage to resist to the temptation to exploit others' contributions [@Hilbig2018]. Lastly, self-control also plays a role, in the sense that more self-control leads to a greater cooperation [@Kocher2017].

In conclusion, values and personality traits have a large impact on human cooperation, and it is thus very important to take them into account when investigating this phenomenon. Interestingly, some social psychological techniques (e.g. elaboration) allow to modify an individual's values [@Sussman2016]. Such techniques could therefore be used to improve cooperation.

@@

observation: majority of papers included are about SD, and very few studies about commons are quantitative --\> future research: quantitative approach to investigate cooperation in common initiatives
